---
title: "STAT615 Group Project Data Download"
author: "Sujata, Carlos, Lesley"
date: "10/12/2020"
output: pdf_document
---

Data Description:

Over 14000 variables [CDC and NY combined]

```{r}
library(tidyverse)
```


```{r}
covid19 <- read.csv("../data/NYT_us-counties.csv")
cdc_svid <- read.csv("../data/SVI2018_US_COUNTY.csv")
```


```{r}
head(cdc_svid)
```


```{r}
head(covid19)
```

```{r}
cdc_svid %>%
    rename("county" = "COUNTY") %>%
    rename("state" = "STATE") %>%
    inner_join(covid19, by=c("county")) -> merged_df

names(merged_df)
dim(merged_df)
```

```{r}
merged_df
write.table(merged_df, file="../data/MergedDF_Oct122020COVID19_CDCSVI_CountyData.tsv", sep = '\t', row.names=FALSE, quote=FALSE)
```

```{r}
# Multiple Linear Regression
# Outcome: COVID19 Deaths
mdl <- lm(deaths ~ population_var1 + income_var2 + age_var3)
```

**Meeting Notes:**

- consider including interaction term
- explain how lasso is relevant to regression analysis and explain connection to ridge regression
  - central idea: explain connection to shrinkage
  
Data Analysis Project Proposal [Due Date TBD ~ Oct 19]

- 1 page description
- List project interests
- List relevant variables
- Possible factors associated with response variable (cummulative COVID19 deaths)
- Predictors being considered
- County level data
- Multiple linear regression
- LASSO for feature selection
- evaluating whether certain variables may have interactions or how they affect outcomes

**Project Update**

Project 1:

- General check in 
- describe what has been done since the inital proposal

Project 2:

- General check in 
- describe what has been done since the inital proposal
Presentation Materials


**Presentation Materials**

Project 1:

- Powerpoint/Slides
- submit as a pdf
- Shiny app
- document accompanying shiny app is needed [pdf format recommended]
- submit as a pdf

Project 2:

- Powerpoint/slides
- submit as a pdf
- Shiny app
- document accompanying shiny app is needed [pdf format recommended]



  
  
  
  
  

TODO: 

Convert state names from all caps to camelcase or vice versa
  + must merge dataframe based on county and state. Some counties may have redundant names

Select features of interest in the merged dataframe
https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/pdf/SVI2018Documentation-H.pdf