---
title: 'STAT615 Regression : Group Project'
author: "Lesley Chapman, Sujata Biradar"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(glmnet) 
library(knitr)
library(fastDummies)
```

```{r}
# Import data
covid19 <- read.csv("../data/NYT_us-counties.csv")
cdc_svid <- read.csv("../data/SVI2018_US_COUNTY.csv")
```

## Data Preprocessing

```{r}
data(swiss)
names(swiss)
x <- model.matrix(Fertility~. , swiss)[,-1]
head(x)
```



```{r}

# Add state abbreviation column
names(state.abb) <- state.name
covid19$ST_ABBR <- state.abb[covid19$state]

# Merge dataframes according to county and state
cdc_svid %>%
    rename("county" = "COUNTY") %>%
    rename("state" = "STATE") %>%
    inner_join(covid19, by=c("county", "ST_ABBR")) %>%
    select(-contains(".x")) %>% 
    select(-contains(".y")) -> merged_df

head(merged_df)
dim(merged_df)
merged_df
# rename df
covid19 <- merged_df
#write.table(merged_df, file="../data/MergedDF.tsv", sep = '\t', row.names=FALSE, quote=FALSE)
```


### Create dataframe with response variable of interestes (confirmed_cases)
```{r}
covid19 %>%
  select(-c(date, fips, deaths, cases, confirmed_deaths, probable_cases, probable_deaths)) -> covid19cases
```

### Convert all categorical variables using dummy_cols package

+ Sujata : add description of why we converted all categorical variables using as.factor() by next meeting or before final presenation; alteraive: we could've usedknitR/fastDummies this however creates more columns and slows compute time

+ Ask: what're the benefits of converting categorical varables to binary as opposed to storing them as factors. We chose to store as factors

```{r, include=FALSE}
str(covid19cases)
```

```{r, include= FALSE}

x <- covid19cases
x$ST_ABBR <- as.factor(x$ST_ABBR)
x$county <- as.factor(x$county)
x$LOCATION <- as.factor(x$LOCATION)
str(x)
```


```{r}
head(x)
```

Dropped any row that had NA in as the response variable
```{r}
covid19cases2 %>% filter(!is.na(confirmed_cases)) -> covid19cases2
```

```{r}
covid19cases2 %>% select(-c(confirmed_cases)) -> covid19cases2_x
x <- model.matrix(ST~., covid19cases2_x) 
y <- covid19cases2$confirmed_cases
```


```{r}
lambda_seq <- 10^seq(2, -2, by = -.1)


set.seed(86)
train = sample(1:nrow(x), nrow(x)/2)
x_test = (-train)
y_test = y[x_test]
 
cv_output <- cv.glmnet(x[train,], 
                       y[train],
                       alpha = 1, 
                       lambda = lambda_seq, 
                       nfolds = 5)
```


```{r}
lambda_seq
```

### Select features selected by LASSO

Coefficients that are 0 should not be included in the model. The following section produces features that we will use within our multiple linear regression
```{r}
best_lam <- cv_output$lambda.min
best_lam

lasso_best <- glmnet(x[train,], y[train], alpha = 1, lambda = best_lam)
lasso_best

CF <- as.matrix(coef(cv_output, cv_output$lambda.1se))
lasso_var <- CF[CF!=0,]

```


```{r}
dim(covid19)
```

```{r}
lasso_vars <- data.frame(lasso_var)
lasso_vars$variables <- rownames(lasso_vars)
lasso_vars <- lasso_vars[!grepl("(Intercept)", lasso_vars$variables),]

# select features suggested by LASSO
covid19cases2_lasso <- covid19cases2[,c("E_UNEMP", "E_NOHSDP", "E_AGE17", "E_SNGPNT", "M_MINRTY", "M_LIMENG", "E_GROUPQ", "EP_MUNIT", "EPL_AGE65", "EPL_AGE17", "EPL_DISABL", "F_MUNIT", "E_UNINSUR", "ST_ABBR_CA", "ST_ABBR_LA", "ST_ABBR_MI", "ST_ABBR_OR", "ST_ABBR_PA", "ST_ABBR_SC", "county_Alameda", "county_Allegheny", "county_Bergen", "county_Bernalillo", "county_Charleston", "county_Clackamas", "county_East Baton Rouge", "county_Erie", "county_Genesee", "county_Greenville", "county_King", "county_Los Angeles", "county_Lubbock", "county_Mecklenburg", "county_Milwaukee", "county_Nassau", "county_Orange", "county_Passaic", "county_Prince George's", "county_Riverside", "county_Santa Clara", "county_Suffolk", "county_Tarrant", "county_Westchester", "county_Woodbury", "county_Yakima", "county_Yuma", "LOCATION_Alameda County, California", "LOCATION_Allegheny County, Pennsylvania", "LOCATION_Bergen County, New Jersey", "LOCATION_Bernalillo County, New Mexico", "LOCATION_Charleston County, South Carolina", "LOCATION_Clackamas County, Oregon", "LOCATION_Clark County, Nevada", "LOCATION_Douglas County, Nebraska", "LOCATION_East Baton Rouge Parish, Louisiana", "LOCATION_Erie County, Pennsylvania", "LOCATION_Essex County, Massachusetts", "LOCATION_Genesee County, Michigan", "LOCATION_Greenville County, South Carolina", "LOCATION_Hall County, Georgia", "LOCATION_Jackson County, Missouri", "LOCATION_King County, Washington", "LOCATION_Lake County, Illinois", "LOCATION_Los Angeles County, California", "LOCATION_Lubbock County, Texas", "LOCATION_Mecklenburg County, North Carolina", "LOCATION_Middlesex County, Massachusetts", "LOCATION_Milwaukee County, Wisconsin", "LOCATION_Nassau County, New York", "LOCATION_Orange County, California", "LOCATION_Passaic County, New Jersey", "LOCATION_Polk County, Iowa", "LOCATION_Prince George's County, Maryland", "LOCATION_Richland County, South Carolina", "LOCATION_Riverside County, California", "LOCATION_Santa Clara County, California", "LOCATION_Suffolk County, New York", "LOCATION_Thurston County, Washington", "LOCATION_Union County, New Jersey", "LOCATION_Westchester County, New York", "confirmed_cases")]
```

### Multiple regression
The following is the initial multiple regression model using the initial set of (LASSO selected) variables
```{r}
covid19_mod <- lm(confirmed_cases~., data=covid19cases2_lasso)
summary(covid19_mod)
```

**Vignettes used for reference**
(Vignette 1)[https://glmnet.stanford.edu/articles/glmnet.html]
(Vignette 2)[https://rstatisticsblog.com/data-science-in-action/machine-learning/lasso-regression/]





